---
title: "Barcelona/LosAngelesUHIStudy"
output: html_notebook
---
Make sure this notebook is located in the assignment folder, with subfolders "LA", "Barcelona" and "Plots".

## Packages
Download and install all necessary packages.
```{r}
list.of.packages <- c("RColorBrewer","classInt", "sp", "rgeos",
                      "sf","rgdal","geojsonio","tidyverse","raster",
                      "rasterVis","RStoolbox","plotly",
                      "htmlwidgets","magrittr","viridis", "tmap", "tmaptools", "shiny", "shinyjs")

#Checks whether any packages needed are not already downloaded&installed.
#Then installs all new packages.
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])] 
if(length(new.packages)) install.packages(new.packages)
```

Library all packages
```{r}
library(sf) #simplefeatures
library(sp)
library(rgeos)
library(rgdal)
library(geojsonio)
library(raster)
library(rasterVis)
library(RStoolbox)
library(plotly)
library(tidyverse)  #prerequisite for lots of packages
library(htmlwidgets)
library(RColorBrewer) #color scales
library(classInt)
library(magrittr) #Pipes - used for plotting
library(viridis)  #color scales
library(tmap)
library(tmaptools)
library(shiny)
library(shinyjs)
```

### Barcelona
Store GeoJSON of Barcelona outline into simple features object. This is used later for cropping the raster.
Also need to change the projection of the barcelona outline to match the projection of raster file
```{r}
barcelonaSF <- st_read("Barcelona/Barcelona.geojson")
newProjection <- "+proj=utm +zone=31 +datum=WGS84 +units=m +no_defs"
barcelonaSF <- st_transform(barcelonaSF, newProjection)
```

## RASTER PRE-PROCESSING
Read in the meta file, and stack the raster bands using the meta file. this is faster than reading in traditionally.
Atmospheric processing, applying DOS2 correction from chavez (1996).
```{r}
barcMTL <- readMeta("Barcelona/May_10_19/LC08_L1TP_198031_20190510_20190521_01_T1_MTL.txt")
espRS <- stackMeta("Barcelona/May_10_19/LC08_L1TP_198031_20190510_20190521_01_T1_MTL.txt")
#determine the SHV on the blue band (B2), with a dark prop of 1% 
hazeDN <- estimateHaze(espRS, hazeBands = "B2_dn", darkProp = 0.01) 
#correct atmospherical scattering to deliver surface reflectance.
atmCorrectedRS <- radCor(espRS, metaData = barcMTL,
                         method = "costz",hazeValues = hazeDN, hazeBands = "B2_dn")
hazeDN
```

Crop the Corrected Raster to the Barcelona extent and border.
```{r}
barcelonaBound <- as(barcelonaSF, Class = "Spatial")
barcelonaExtent <- extent(barcelonaBound)
barcelonaRSCrop <- crop(atmCorrectedRS, barcelonaExtent)
barcelonaRSMask <- mask(barcelonaRSCrop, barcelonaBound)
```

## PROCESSING - SAVI, IBI, LST, UHI, UHS, UTFVI
Create Index Function.
```{r}
indexFunction <- function(band1, band2) {
  Index <- (band1 - band2) / (band1 + band2)
  return(Index)
}
```
SAVI
used to detect vegetation over NDVI because more sensitive for areas where vegetation density is low (i.e cities)
L = Constant that depends on amount of vegetation,
0 = highest vegetation density, 1 = lowest vegetation density.
Generally taken as 0.5.
```{r}
L = 0.5
barcelonaSAVI <- ((barcelonaRSMask$B5_sre - barcelonaRSMask$B4_sre)*(1+L))/
                    (barcelonaRSMask$B5_sre + barcelonaRSMask$B4_sre + L)
```
MNDWI
used to extract water from landsat8, more sensitive to areas such as cities where water is not commonly found.
```{r}
barcelonaMNDWI <- indexFunction(barcelonaRSMask$B3_sre,barcelonaRSMask$B6_sre)
```

IBI
We can now use both of these indexes to just extract built land using IBI.
Need to add one to all indices to remove negative numbers for calculation first.
```{r}
NDBI1 <- barcelonaNDBI + 1
SAVI1 <- barcelonaSAVI + 1
MNDWI1 <- barcelonaMNDWI + 1
barcelonaIBI <- ((NDBI1-((SAVI1+MNDWI1)/2))/
                   (NDBI1+((SAVI1+MNDWI1)/2)))
```

##LST
As we have performed atmospheric correction, we already have the Brightness temperature values in bands 10.#So no need to recalculate.
```{r}
barcBT1 <- barcelonaRSMask$B10_bt

```

Next calculate fractional vegetation value using SAVI values. This used to calculate emissivity. 

```{r}
barcFracV=(barcelonaSAVI-0.2/0.5-0.2)^2
emiss1=0.004*barcFracV+0.986
```

p calculation and define lambda
```{r}
Boltzmann=1.38*10e-23
Plank=6.626*10e-34
c=2.998*10e8
p=Plank*(c/Boltzmann)
lambda=1.09e-5

```
Finally calculate LST equation and convert to celsius
```{r}
barcLST <-barcBT1/(1 +(lambda*barcBT1/p)*log(emiss1))
barcLST<-barcLST-273.15
```
## Determining UHI, non-UHI, UHS and UTFVI
```{r}
sd1 <- cellStats(barcLST, stat = "sd")
mean1 <- cellStats(barcLST, stat = "mean")
barcUHI <- barcLST > (mean1 + (0.5*sd1))     #(Ma,Kuang & Huang 2010)
barcNonUHI <- barcLST <= (mean1 + (0.5*sd1)) #(Guha,Govil & Mukherjee 2017)
barcUHS <- barcLST > (mean1 + (2*sd1))
barcUTFVI <- (barcLST - mean1)/mean1
```

## LOS ANGELES
Same process as barcelona, so putting all in one chunk (May take 10 mins or so to run.)

```{r}
laSF <- st_read("LA/LosAngeles.geojson")

laMTL <- readMeta("LA/May_30_19/LC08_L1TP_041036_20190530_20190605_01_T1_MTL.txt")
laRS <- stackMeta("LA/May_30_19/LC08_L1TP_041036_20190530_20190605_01_T1_MTL.txt")
newProjection <- "+proj=utm +zone=11 +datum=WGS84 +units=m +no_defs"
laSF <- st_transform(laSF, newProjection)
hazeDN1 <- estimateHaze(laRS, hazeBands = "B2_dn", darkProp = 0.01)
laCorRS <- radCor(laRS, metaData = laMTL,
                         method = "costz",hazeValues = hazeDN1, hazeBands = "B2_dn")

laBound <- as(laSF, Class = "Spatial")
laExtent <- extent(laBound)
laRSCrop <- crop(laCorRS, laExtent)
laRSMask <- mask(laRSCrop, laBound)

L = 0.5
laSAVI <- ((laRSMask$B5_sre - laRSMask$B4_sre)*(1+L))/
                    (laRSMask$B5_sre + laRSMask$B4_sre + L)
laNDBI <- indexFunction(laRSMask$B6_sre, laRSMask$B5_sre)
laMNDWI <- indexFunction(laRSMask$B3_sre,laRSMask$B6_sre)
NDBI2 <- laNDBI + 1
SAVI2 <- laSAVI + 1
MNDWI2 <- laMNDWI + 1
laIBI <- (NDBI2-(SAVI2+MNDWI2)/2)/
            (NDBI2+(SAVI2+MNDWI2)/2)

laBT <- laRSMask$B10_bt
laFracV=(laSAVI-0.2/0.5-0.2)^2
emiss2=0.004*laFracV+0.986
Boltzmann=1.38*10e-23
Plank=6.626*10e-34
c=2.998*10e8
p=Plank*(c/Boltzmann)
lambda=1.09e-5
laLST <-laBT/(1 +(lambda*laBT/p)*log(emiss2))
laLST<-laLST-273.15

sd2 <- cellStats(laLST, stat = "sd")
mean2 <- cellStats(laLST, stat = "mean")
laUHI <- laLST > (mean2 + (0.5*sd2))      
laNonUHI <- laLST <= (mean2 + (0.5*sd2))  
laUHS <- laLST > (mean2 + (2*sd2))  
laUTFVI <- (laLST - mean2)/mean2

```

## PLOTTING
Create map functions for plotting barcelona and LA, so that dont need to rewrite everytime.
Store colour palettes and SAVI breaks too.
```{r}
createMap <- function(ras, geoshp, colour,br, legTitle, Mtitle){
  tm_shape(ras) + 
    tm_raster(style = "fixed", title = Mtitle,
              
              palette = colour, midpoint = NA, breaks = br)+
    tm_legend(outside = FALSE, legend.position = c("right", "bottom"))+
    tm_shape(geoshp) + 
    tm_borders(alpha =  0.4, col = "black") + 
    tm_layout(inner.margins = c(0.1, 0.1, 0.1, 0.1),
              title = legTitle,
              title.size = 1.1,
              title.position = c("center", "top")) +
    tm_compass(type = "4star", size = 3, text.size = 1,position = c("left", "top")) +
    tm_scale_bar(position = c("left","bottom"))
  
}
createMap2 <- function(ras, geoshp, colour,br, legTitle, Mtitle){
  tm_shape(ras) + 
    tm_raster(style = "fixed", title = Mtitle,
              
              palette = colour, midpoint = NA, breaks = br)+
    tm_legend(outside = FALSE, legend.position = c("left", "bottom"))+
    tm_shape(geoshp) + 
    tm_borders(alpha =  0.4, col = "black") + 
    tm_layout(inner.margins = c(0.08, 0.1, 0.1, 0.1),
              title = legTitle,
              title.size = 1.1,
              title.position = c("center", "top")) +
    tm_compass(type = "4star", size = 3, text.size = 1,position = c("left", "top")) +
    tm_scale_bar(position = c("right","bottom"))
  
}
#Colour palettes
RedGreen <- c("#A93226","#E67E22","#F1C40F","#52BE80","#196F3D")
GreenRed <- c("#196F3D","#52BE80","#F1C40F","#E67E22","#A93226")
#Break points for SAVI threshold intervals
cuts = c(-1,0,0.1, 0.2, 0.3,1)
```

Create Maps for LST, IBI, SAVI distributions
```{r}
SAVImap1 <- createMap(barcelonaSAVI, barcelonaSF, RedGreen, cuts, "Barcelona - SAVI","SAVI")
SAVImap2 <- createMap2(laSAVI, laSF, RedGreen, cuts, "Los Angeles - SAVI", "SAVI")

SAVIplot <- tmap_arrange(SAVImap1, SAVImap2, ncol = 2)
SAVIplot

IBImap1 <- createMap(barcelonaIBI, barcelonaSF, GreenRed, c(-1,-0.2,-0.1,0,0.1,1), "Barcelona - IBI", "IBI")
IBImap2 <- createMap2(laIBI, laSF, GreenRed, c(-1,-0.2,-0.1,0,0.1,1), "Los Angeles - IBI", "IBI")
IBIplot <- tmap_arrange(IBImap1, IBImap2, ncol=2)
IBIplot
LSTmap1 <- createMap(barcLST, barcelonaSF, GreenRed, c(14,22, 24, 25, 26,30, 36),"Barcelona - LST", "LST (Celsius)")
LSTmap2 <- createMap2(laLST, laSF, GreenRed, c(14,22,30,33,36,40,48), "Los Angeles - LST", "LST (Celsius)")
LSTmap <- tmap_arrange(LSTmap1, LSTmap2, ncol=2)
LSTmap
```

Crop LST to UHI map to work out midpoints for next plot.
```{r}
bUHIcrop <- mask(barcLST, barcUHI, maskvalue = 0, updatevalue = NA)
bNUHIcrop <- mask(barcLST,barcUHI, maskvalue = 1, updatevalue = NA)
lUHIcrop <- mask(laLST, laUHI, maskvalue = 0, updatevalue = NA)
lNUHIcrop <- mask(laLST, laUHI, maskvalue = 1, updatevalue = NA)
bUHScrop <- mask(barcLST, barcUHS, maskvalue = 0, updatevalue = NA)
lUHScrop <- mask(laLST, laUHS, maskvalue = 0, updatevalue = NA)
#Simply find minimum value of bUHIcrop and lUHIcrop to use as midpoints in next plot.
UHImap1 <- tm_shape(barcLST) +
  tm_raster(style = "cont", title = "LST (Celsius)", midpoint = 25.764, palette = "-RdYlGn",n = 9, contrast = c(0.4, 1)) +
  tm_legend(outside = FALSE, legend.position = c("right", "bottom"))+
  tm_shape(barcelonaSF) + 
  tm_borders(alpha =  0.4, col = "black") + 
  tm_layout(inner.margins = c(0.05, 0.02, 0.13, 0.1),
            title = "Barcelona LST\n UHI & Non-UHI",
            title.size = 1.1,
            title.position = c("center", "top")) +
  tm_compass(type = "4star", size = 2.5, text.size = 0.8,position = c("left", "top")) +
  tm_scale_bar(position = c("left","bottom"))


UHImap2 <- tm_shape(laLST) +
  tm_raster(style = "cont", title = "LST (Celsius)", midpoint = 35.51689, palette = "-RdYlGn",n = 9, contrast = c(0.4, 1)) +
  tm_legend(outside = FALSE, legend.position = c("left", "bottom"))+
  tm_shape(laSF) + 
  tm_borders(alpha =  0.4, col = "black") + 
  tm_layout(inner.margins = c(0.08, 0.1, 0.13, 0.1),
            title = "Los Angeles LST\n UHI & Non-UHI",
            title.size = 1.1,
            title.position = c("center", "top")) +
  tm_compass(type = "4star", size = 2.5, text.size = 0.8,position = c("left", "top")) +
  tm_scale_bar(position = c("right","bottom"))

UHIplot <- tmap_arrange(UHImap1, UHImap2, ncol = 2)
UHIplot
```

UTFVI and UHS plots
```{r}
UTFVIbins <- c(-1,0,0.005,0.01,0.015,0.02,1)
UTFVImap1 <- createMap(barcUTFVI, barcelonaSF, GreenRed, UTFVIbins, "Barcelona - UTFVI", "UTFVI")
UTFVImap2 <- createMap2(laUTFVI, laSF, GreenRed, UTFVIbins, "Los Angeles - UTFVI", "UTFVI")
UTFVIplot <- tmap_arrange(UTFVImap1, UTFVImap2, ncol = 2)
UTFVIplot

UHSmap1 <- tm_shape(barcUHS) +
  tm_raster(style = 'cont',title = "Legend" ,palette = c("white", "red"), labels = c("Barcelona","Urban Hot Spots"))+
  tm_legend(outside = F, legend.position = c("right", "bottom"), frame = "black")+
  tm_shape(barcelonaSF) +
  tm_borders(alpha = 0.4, col = "black") +
  tm_layout(inner.margins = c(0.05, 0.02, 0.13, 0.1), 
            title = "Barcelona - Urban Heat Spots (UHS)",
            title.size = 1.1,
            title.position = c("center", "top")) +
  tm_compass(type = "4star", size = 2.5, text.size = 0.8,position = c("left", "top")) +
  tm_scale_bar(position = c("left","bottom"))

UHSmap2 <- tm_shape(laUHS) +
  tm_raster(style = 'cont',title = "Legend" ,palette = c("white", "red"), labels = c("Los Angeles","Urban Hot Spots"))+
  tm_legend(outside = F, legend.position = c("right", "bottom"), frame = "black")+
  tm_shape(laSF) +
  tm_borders(alpha = 0.4, col = "black") +
  tm_layout(inner.margins = c(0.05, 0.2, 0.13, 0.2), 
            title = "Los Angeles - Urban Heat Spots (UHS)",
            title.size = 1.1,
            title.position = c("center", "top")) +
  tm_compass(type = "4star", size = 2.5, text.size = 0.8,position = c("left", "top")) +
  tm_scale_bar(position = c("left","bottom"))
UHSmap1
UHSmap2

```

Work out the area % of UTFVI
```{r}
utfviDF1 <- as.data.frame(na.omit(values(barcUTFVI)))
utfviDF2 <- as.data.frame(na.omit(values(laUTFVI)))

bclass1 <- sum(utfviDF1[] < 0)
bclass2 <- sum(utfviDF1[] >= 0 & utfviDF1[] < 0.005)
bclass3 <- sum(utfviDF1[] >= 0.005 & utfviDF1[] < 0.01)
bclass4 <- sum(utfviDF1[] >= 0.01 & utfviDF1[] < 0.015)
bclass5 <- sum(utfviDF1[] >= 0.015 & utfviDF1[] < 0.02)
bclass6 <- sum(utfviDF1[] >= 0.02)

bTot <- bclass1+bclass2+bclass3+bclass4+bclass5+bclass6
b1 <- (bclass1/bTot)*100
b2 <- (bclass2/bTot)*100
b3 <- (bclass3/bTot)*100
b4 <- (bclass4/bTot)*100
b5 <- (bclass5/bTot)*100
b6 <- (bclass6/bTot)*100

lclass1 <- sum(utfviDF2[] < 0)
lclass2 <- sum(utfviDF2[] >= 0 & utfviDF2[] < 0.005)
lclass3 <- sum(utfviDF2[] >= 0.005 & utfviDF2[] < 0.01)
lclass4 <- sum(utfviDF2[] >= 0.01 & utfviDF2[] < 0.015)
lclass5 <- sum(utfviDF2[] >= 0.015 & utfviDF2[] < 0.02)
lclass6 <- sum(utfviDF2[] >= 0.02)

lTot <- lclass1+lclass2+lclass3+lclass4+lclass5+lclass6
l1 <- (lclass1/lTot)*100
l2 <- (lclass2/lTot)*100
l3 <- (lclass3/lTot)*100
l4 <- (lclass4/lTot)*100
l5 <- (lclass5/lTot)*100
l6 <- (lclass6/lTot)*100

b1
b2
b3
b4
b5
b6
l1
l2
l3
l4
l5
l6
```
## Linear relationships
stack all data that you want to explore the relationship of, and then convert to dataframe.
remove null values too.
```{r}
barcStack <- stack(barcLST, barcelonaSAVI, barcelonaIBI)
barcDF <- as.data.frame(na.omit(values(barcStack)))
laStack <- stack(laLST, laSAVI, laIBI)
laDF <- as.data.frame(na.omit(values(laStack)))
#barcelona
bTempVeg <- lm(layer.2~layer.1, data = barcDF)
summary(bTempVeg)
bTempIBI <- lm(layer.3 ~ layer.1, data = barcDF)
summary(bTempIBI)
#LA
lTempVeg <- lm(layer.2~layer.1, data = laDF)
summary(lTempVeg)
lTempIBI <- lm(layer.3 ~ layer.1, data = laDF)
summary(lTempIBI)
```
## Write out plots
```{r}
tmap_save(SAVIplot, filename = "Plots/SAVIplot.png")
tmap_save(UHIplot, filename = "Plots/UHIplot.png")
tmap_save(UHSmap1, filename = "Plots/UHSbarcelona.png")
tmap_save(UHSmap2, filename = "Plots/UHSla.png")
tmap_save(IBIplot, filename = "Plots/IBIplot.png")
tmap_save(LSTmap, filename = "Plots/LSTplot.png")
tmap_save(UTFVIplot, filename = "Plots/UTFVIplot.png")
```

